Adaptive Sequential Model Selection

### 2015-05-06 paper.tex
defined "generalized selected-model p-values"

% latex table generated in R 3.0.2 by xtable 1.7-1 package
% Mon May  4 21:11:15 2015
\begin{table}[ht]
  \centering
  \begin{tabular}{l|ccccc|c}
    \multicolumn{7}{c}{Saturated-Model $p$-Values 
      (\% of $10^6$ Simulations)}\\
    \hline
    \multicolumn{7}{c}{$p_2(Y)$}\\[5pt]
    ${\large p_1(Y)}$ & (0,0.2] & (0.2,0.4] & (0.4,0.6] & (0.6,0.8] & (0.8,1] & \textbf{Total} \\ 
    \hline
    (0,0.2] & 1.0 & 2.7 & 4.2 & 5.6 & 6.7 & 20.1 \\ 
    (0.2,0.4] & 1.4 & 3.4 & 4.5 & 5.2 & 5.5 & 20.0 \\ 
    (0.4,0.6] & 2.3 & 4.3 & 4.7 & 4.5 & 4.2 & 20.0 \\ 
    (0.6,0.8] & 4.2 & 5.4 & 4.3 & 3.4 & 2.7 & 20.0 \\ 
    (0.8,1] & 11.1 & 4.3 & 2.3 & 1.4 & 1.0 & 20.0 \\ 
    \hline
    \textbf{Total} & 19.9 & 20.0 & 20.0 & 20.1 & 20.0 & 100.0 \\ 
    \hline
  \end{tabular}
  \caption{Two-way contingency table of saturated-model $p$-values $(p_1(Y), p_2(Y))$ for Example~\ref{ex:bivariate}, after binning into cells of height and width 0.2. We report the percentage of $p$-value pairs falling into each cell out of one million simulations from the global null hypothesis, $\mu=0$. Both $p$-values are marginally uniform but strongly dependent, with a correlation of $-48\%$.}
\label{tab:bv_twoWayTable}
\end{table}

@@

\subsection{Sequential Sufficiency Principles}\label{sec:sufficiency}

\subsubsection{$p$-Values and the Sufficiency Principle}\label{sec:pvalSP}

A $p$-value for testing model $M_{k-1}$ against $M_k$ satisfies the (conditional) {\em sufficiency principle} if, on the event $\{M_{k-1}=m_{k-1}, M_k=m_k\}$, it depends only on sufficient statistics of $M_k$.

Based on the law
\[
\L_{m_{k-1}}\left(T_k \;\;\mid\;\; T_{k-1},\; M_{k-1}=m_{k-1},\; M_k=m_k\right)
\]

If $T(Y)$ is a complete sufficient statistic, then any exact $p$-value must have the latter property.

@@

\WFcomment{There is a filtration interpretation when you have the appropriate sufficiency properties.} Let $\sF_{k,\ell}$ denote the $\sigma$-algebra generated by $M_{[k]}$ and $p_{[\ell]}$.
\begin{align*}
  \sF_{k,\ell} &= \sF(M_{[k]},p_{[\ell]})\\
  \sF_0 &\underlabel_{\text{selection } 1} \sF_{1,0} \underlabel_{\text{inference } 1}
  \sF_{1,1} \;\;\sub \cdots \sub\;\;
  \sF_{d-1,d-1} \underlabel_{\text{selection } d} \sF_{d,d-1}
  \underlabel_{\text{inference } d} \sF_{d,d}
\end{align*}

>>>>

% latex table generated in R 3.0.2 by xtable 1.7-1 package
% Mon May  4 21:11:15 2015
\begin{table}[ht]
  \centering
  \begin{tabular}{l|ccccc|c}
    \multicolumn{7}{c}{Saturated-Model $p$-Values 
      (\% of $10^6$ Simulations)}\\[7pt]
    \hline
    \multicolumn{7}{c}{}\\[-1.5ex]
    \multicolumn{7}{c}{$p_2(Y)$}\\[5pt]
    ${\large p_1(Y)}$ & (0,0.2] & (0.2,0.4] & (0.4,0.6] & (0.6,0.8] & (0.8,1] & \textbf{Total} \\ 
    \hline
    (0,0.2] & 1.0 & 2.7 & 4.2 & 5.6 & 6.7 & 20.1 \\ 
    (0.2,0.4] & 1.4 & 3.4 & 4.5 & 5.2 & 5.5 & 20.0 \\ 
    (0.4,0.6] & 2.3 & 4.3 & 4.7 & 4.5 & 4.2 & 20.0 \\ 
    (0.6,0.8] & 4.2 & 5.4 & 4.3 & 3.4 & 2.7 & 20.0 \\ 
    (0.8,1] & 11.1 & 4.3 & 2.3 & 1.4 & 1.0 & 20.0 \\ 
    \hline
    \textbf{Total} & 19.9 & 20.0 & 20.0 & 20.1 & 20.0 & 100.0 \\ 
    \hline
  \end{tabular}
  \caption{Two-way contingency table of saturated-model $p$-values $(p_1(Y), p_2(Y))$ for Example~\ref{ex:bivariate}, after binning into cells of height and width 0.2. We report the percentage of $p$-value pairs falling into each cell out of one million simulations from the global null hypothesis, $\mu=0$. Both $p$-values are marginally uniform but strongly dependent, with a correlation of $-48\%$.}
\label{tab:bv_twoWayTable}
\end{table}

@@

\subsection{Conditions for Independent $p$-Values}\label{sec:sufficiency}

To obtain a sequence of $p$-values, we must specify what model selection algorithm $M_{[d]}$ to use, as well as how to construct $p$-values at each step. In this section we consider what conditions are sufficient to guarantee that the resulting $p$-value sequence is independent. Sections~\ref{sec:pvalSP}--\ref{sec:modelSPSP} discuss sufficient conditions on the $p$-values and the selection algorithm under which single-step $p$-values are automatically independent. Section~\ref{sec:selectionVariables} discusses how we can create independent $p$-values by conditioning on finer selection variables at each step.

Essentially, we will want to partition the information in $Y$ according to the filtration:
\begin{align}\nonumber
  \sF(M_0,T_0) &\underlabel_{\text{selection } 1} 
  \sF(M_{[1]},T_0) \underlabel_{\text{inference } 1}
  \sF(M_{[1]},T_1) \quad \sub \;\;\cdots\\[8pt]
  \label{eq:infoPartition}
  \cdots\;\; \sub \quad&
  \sF(M_{[d-1]},T_{d-1}) \underlabel_{\text{selection } d}
  \sF(M_{[d]},T_{d-1}) 
  \underlabel_{\text{inference } d}
  \sF(M_{[d]}, T_d)
\end{align}

\subsubsection{Generalized Selected-Model $p$-Values}\label{sec:pvalSP}

For a candidate pair $(m_{k-1},m_k)$ of models to test at step $k$, define the selection event $A = \{M_{k-1},M_k\}$. We say $p(Y)$ is a {\em generalized selected-model $p$-value} for testing $m_{k-1}$ against $m_k$ at step $k$ if it is ``based on'' the law
\[
\L_{m_{k-1}}\left(T_k \;\;\mid\;\; T_{k-1},\; M_{k-1}=m_{k-1},\;M_k=m_k\right)
\]
for some minimal sufficient statistics $T_{k-1}(Y)$ and $T_k(Y)$ of $m_{k-1}$ and $m_k$, respectively. More precisely, $p(Y)$ must be uniform or conservative conditional on $T_{k-1}(Y)$, and it must be a function of $T_k(Y)$ on $A$.

@@

\begin{comment}
\WFcomment{There is a filtration interpretation when you have the appropriate sufficiency properties.} Let $\sF_{k,\ell}$ denote the $\sigma$-algebra generated by $M_{[k]}$ and $p_{[\ell]}$.
\begin{align*}
  \sF_{k,\ell} &= \sF(M_{[k]},p_{[\ell]})\\
  \sF_0 &\underlabel_{\text{selection } 1} \sF_{1,0} \underlabel_{\text{inference } 1}
  \sF_{1,1} \;\;\sub \cdots \sub\;\;
  \sF_{d-1,d-1} \underlabel_{\text{selection } d} \sF_{d,d-1}
  \underlabel_{\text{inference } d} \sF_{d,d}
\end{align*}
\end{comment}
